OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-06-23 paper/paper.Rmd
Rename results figure slug

While _de novo_ clustering produces high quality OTUs, researchers may prefer to
perform reference clustering to a public database because reference-based
methods produce consistent OTUs and are generally faster than _de novo_ methods.
In closed reference mode, sequences that cannot be assigned to reference OTUs
are thrown out, so that the final clustering contains only OTUs that exist in
the reference.
To test how OptiFit performs for this purpose, we fit each dataset to three
databases of reference OTUs: the Greengenes database, the SILVA non-redundant
database, and the Ribosomal Database Project (RDP) [@desantis_greengenes_2006;
@quast_silva_2013; @cole_ribosomal_2014].
Reference OTUs for each database were created by performing _de novo_ clustering
with OptiClust at a distance threshold of 3% (see Figure \@ref(fig:workflow)).
The _de novo_ MCC scores were `r dn_dbs_mcc` for `r names(dn_dbs_mcc)`, 
respectively.
Fitting sequences to Greengenes and SILVA in closed reference mode performed
similarly, with median MCC scores of `r closed_fit_gg_mcc` and 
`r closed_fit_silva_mcc` respectively, while the median MCC 
dropped to `r closed_fit_rdp_mcc` when fitting to RDP 
(see Figure \@ref(fig:workflow)).
For comparison, clustering datasets _de novo_ with
OptiClust produced an average MCC score of `r opticlust_mcc`.
This gap in OTU quality mostly disappeared when clustering in open reference
mode, which produced median MCCs of `r open_fit_gg_mcc` with greengenes, 
`r open_fit_silva_mcc` with SILVA, and `r open_fit_rdp_mcc` with RDP.
Thus, open reference OptiFit produced OTUs of very similar quality as _de novo_
clustering, and closed reference OptiFit followed closely behind as long as a
suitable reference database was chosen.

```{r workflow, dpi=300, fig.dim=c(4,3), fig.cap="The Analysis Workflow"}
# TODO: write caption
here('figures', 'workflow.tiff') %>% 
  tiff::readTIFF() %>% 
  grid::grid.raster() 
```

@@

```{r results, dpi=100, fig.dim=c(6,6), fig.cap="Benchmarking results"}
# TODO: write caption
here('figures', 'results.tiff') %>% 
  tiff::readTIFF() %>% 
  grid::grid.raster() 
```

>>>>

While _de novo_ clustering produces high quality OTUs, researchers may prefer to
perform reference clustering to a public database because reference-based
methods produce consistent OTUs and are generally faster than _de novo_ methods.
In closed reference mode, sequences that cannot be assigned to reference OTUs
are thrown out, so that the final clustering contains only OTUs that exist in
the reference.
To test how OptiFit performs for this purpose, we fit each dataset to three
databases of reference OTUs: the Greengenes database, the SILVA non-redundant
database, and the Ribosomal Database Project (RDP) [@desantis_greengenes_2006;
@quast_silva_2013; @cole_ribosomal_2014].
Reference OTUs for each database were created by performing _de novo_ clustering
with OptiClust at a distance threshold of 3% (see Figure \@ref(fig:workflow)).
The _de novo_ MCC scores were `r dn_dbs_mcc` for `r names(dn_dbs_mcc)`, 
respectively.
Fitting sequences to Greengenes and SILVA in closed reference mode performed
similarly, with median MCC scores of `r closed_fit_gg_mcc` and 
`r closed_fit_silva_mcc` respectively, while the median MCC 
dropped to `r closed_fit_rdp_mcc` when fitting to RDP 
(see Figure \@ref(fig:results-sum)).
For comparison, clustering datasets _de novo_ with
OptiClust produced an average MCC score of `r opticlust_mcc`.
This gap in OTU quality mostly disappeared when clustering in open reference
mode, which produced median MCCs of `r open_fit_gg_mcc` with greengenes, 
`r open_fit_silva_mcc` with SILVA, and `r open_fit_rdp_mcc` with RDP.
Thus, open reference OptiFit produced OTUs of very similar quality as _de novo_
clustering, and closed reference OptiFit followed closely behind as long as a
suitable reference database was chosen.

```{r workflow, dpi=300, fig.dim=c(4.5,3.5), fig.cap="**The Analysis Workflow.**"}
# TODO: write caption
here('figures', 'workflow.tiff') %>% 
  tiff::readTIFF() %>% 
  grid::grid.raster() 
```

@@

```{r results-sum, dpi=100, fig.dim=c(6,6), fig.cap="**Benchmarking results**"}
# TODO: write caption
here('figures', 'results-sum.tiff') %>% 
  tiff::readTIFF() %>% 
  grid::grid.raster() 
```
